GVN-Hoist: Hoisting Computations from Branches

### 2016-08-11 hoist.tex
add related work, move ILP to cost model

The main contributions of this paper are:
\begin{itemize}
\item a new algorithm to hoist computations from branches
\item cost model to reduce live-range and reduce spills
\item performance evaluation of the implementation in LLVM
\item comparison against other algorithms for code hoisting
\end{itemize}

@@

graph algorithms \cite{tarjan},
SSA \cite{cytron}
GVN \cite{rosen1988global,click1995global}
CFG, DOM \cite{dragonbook}
SESE, SEME \cite{sese}
MemorySSA \cite{novillo2007memory}

\subsection{Global Value Numbering}

Code hoisting can also reduce the critical path length of execution in out of
order machines. As more instructios are available at the hoisting point, the
hardware has more instructions to reorder. Following example illustrates how
hoisting can improve performance by exposing more ILP.

@@

\section{Code hoisting}
For the code hoisting we follow a very systematic approach to writing a compiler
optimization pass. As such this pass can be divided into the following five
steps:

@@

>>>>

The main contributions of this paper are:
\begin{itemize}
\item a new algorithm to hoist computations from branches,
\item cost model to reduce live-range and reduce spills,
\item performance evaluation of the implementation in LLVM,
\item comparison against other algorithms for code hoisting.
\end{itemize}

@@

@@

\section{Code hoisting}

\subsection{}
The algorithm for code hoisting uses several common representations of the
program that we shortly describe below:
\begin{itemize}
\item Control Flow Graph (CFG) and the Dominance (DOM) and Post-Dominance (PDOM)
  relations \cite{dragonbook};
\item Single Entry Single Exit (SESE) \cite{sese} and Single Entry Multiple Exit
  (SEME) regions;
\item Static Single Assignment (SSA) \cite{cytron};
\item Global Value Numbering (GVN) \cite{rosen1988global,click1995global}: to
  identify similar computations compilers use GVN.  Each expression is given a
  unique number and the expressions that the compiler can prove identical are
  given the same number;
\item Memory SSA \cite{novillo2007memory}: memory operations that the compiler
  is able to prove in dependence are linked through use-def chains;
\end{itemize}

For the code hoisting we follow a very systematic approach to writing a compiler
optimization pass. As such this pass can be divided into the following five
steps:

@@

\subsubsection{Exposing more ILP}
Code hoisting can also reduce the critical path length of execution in out of
order machines. As more instructios are available at the hoisting point, the
hardware has more instructions to reorder. Following example illustrates how
hoisting can improve performance by exposing more ILP.
